The dictionary for characters of smooth group schemes

### 2016-05-21 followup.tex
Further changes to introduction.

The first aim of this paper is to repair this defect in the function-sheaf dictionary
 by describing a full subcategory $\CCS(G)$ of $\CS(G)$ so that $\TrFrob{G}$ restricts to an isomorphism $\CCSiso{G} \to G(k)^*$.
We refer to the resulting objects as \emph{commutative character sheaves}, since the passage from $\CS(G)$ to $\CCS(G)$ involves a condition that exchanges the inputs to the multiplication morphism on $G$ (see Definition \ref{def:CCScom}).  
When $G$ is connected, all character sheaves on $G$ are commutative.
This category clarifies several questions about character sheaves on $G$.  Invisible character sheaves \cite{cunningham-roe:13a}*{Def. 2.8} are precisely those $\cs{L}$ with $\TrFrob{G}(\cs{L}) = 1$ that are not commutative.  Moreover, $\TrFrob{G}^{-1} : G(k)^* \to \CCSiso{G}$ provides a canonical splitting of $\TrFrob{G} : \CSiso{G} \to G(k)^*$ \cite{cunningham-roe:13a}*{Rem. 3.7}.

@@

In order to further justify referring to objects in $\CCS(G)$ as commutative character sheaves,
suppose for the moment that $G$ is a connected, reductive algebraic group over $\Fq$.
In this case, every object in $\CCS(G)$ is a Frobenius-stable character sheaf on $G$ in the sense of \cite{lusztig:85a}*{Def.~2.10}.
Moreover, if $A$ is a character sheaf on $G$ in this sense, and if the characteristic function of $A$ is a character of $G(k)$, then the dictionary determines an object in $\CCS(G)$ from $A$.
These two rules determine an injection from $\CCSiso{G}$ to isomorphism classes of Frobenius-stable character sheaves on $G$.
Accordingly, we may refer to character sheaves on $G$ that arise in this manner as commutative character sheaves.
Of course, such sheaves represent a small part of Lusztig's geometrization of characters of representations of connected, reductive groups over finite fields.

@@

We now summarize the sections of the paper in more detail.
%
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes sense when $G$ is not commutative.  
%
We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
%
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our second main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which require a development of equivariant linear character sheaves.

In Section~\ref{sec:types} we give an application of Theorem~\ref{thm:geo} by geometrizing types for supercuspidal representations of $p$-adic groups. The proof of this statement, which is made precise in Theorem~\ref{thm:geotypes}, is given in Section~\ref{ssec:geotypes}. 
 %Specifically, we show that if $C$ is a compact $p$-adic group and $\rho$ is a representation of $C$ and $(C,\rho)$ is a type for a tamely ramified $p$-adic group then there is a  group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a virtual Weil sheaf $\mathcal{A}$ on $G$ such that $\trFrob{\cs{A}} = \trace(\rho)$.
As preparation for the proof, we review some facts about the Heisenberg-Weil representation and its geometrization, in Section~\ref{ssec:Jacobi}.
Then, in Section~\ref{ssec:review}, we review Yu's theory of types and his study of smooth integral models, in Section~\ref{ssec:models}.  
These elements are pulled together in Section~\ref{ssec:geotypes}, where the proof Theorem~\ref{thm:geotypes} is given.

>>>>

The first aim of this paper is to repair this defect in the function-sheaf dictionary
 by describing a full subcategory $\CCS(G)$ of $\CS(G)$ so that $\TrFrob{G}$ restricts to an isomorphism $\CCSiso{G} \to G(k)^*$.
We refer to the resulting objects as \emph{commutative character sheaves}, since the passage from $\CS(G)$ to $\CCS(G)$ involves a condition that exchanges the inputs to the multiplication morphism on $G$ (see Definition \ref{def:CCScom}).  
When $G$ is connected, all character sheaves on $G$ are commutative.
This category clarifies several questions about $\CS(G)$. \todo{Small change here.} Invisible character sheaves \cite{cunningham-roe:13a}*{Def. 2.8} are precisely those $\cs{L}$ with $\TrFrob{G}(\cs{L}) = 1$ that are not commutative.  Moreover, $\TrFrob{G}^{-1} : G(k)^* \to \CCSiso{G}$ provides a canonical splitting of $\TrFrob{G} : \CSiso{G} \to G(k)^*$ \cite{cunningham-roe:13a}*{Rem. 3.7}.

@@

@@

In order to provide further justification for referring to objects in $\CCS(G)$ as commutative character sheaves, suppose for the moment that $G$ is a connected, reductive algebraic group over $\Fq$.
In this case, it can be shown that there is a forgetful functor taking every object in $\CCS(G)$ to a Frobenius-stable character sheaf on $G$, in the sense of \cite{lusztig:85a}*{Def.~2.10}, for which the characteristic function is a character of $G(k)$.
Of course, such sheaves represent a small part of Lusztig's geometrization of characters of representations of connected, reductive groups over finite fields,
but they are precisely the part needed to describe characters of such groups.
\bigskip

We now summarize the sections of the paper in more detail.
%
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes sense when $G$ is not commutative.  
%
We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
%
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our second main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which require a development of equivariant linear character sheaves. \todo{There was a paragraph break here. Why?}
% 
In Section~\ref{sec:types} we give an application of Theorem~\ref{thm:geo} by geometrizing types for supercuspidal representations of $p$-adic groups. The proof of this statement, which is made precise in Theorem~\ref{thm:geotypes}, is given in Section~\ref{ssec:geotypes}. 
 %Specifically, we show that if $C$ is a compact $p$-adic group and $\rho$ is a representation of $C$ and $(C,\rho)$ is a type for a tamely ramified $p$-adic group then there is a  group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a virtual Weil sheaf $\mathcal{A}$ on $G$ such that $\trFrob{\cs{A}} = \trace(\rho)$.
As preparation for the proof, we review some facts about the Heisenberg-Weil representation and its geometrization, in Section~\ref{ssec:Jacobi}.
Then, in Section~\ref{ssec:review}, we review Yu's theory of types and his study of smooth integral models, in Section~\ref{ssec:models}.  
These elements are pulled together in Section~\ref{ssec:geotypes}, where the proof Theorem~\ref{thm:geotypes} is given.
\bigskip
